---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activity <- read.csv(file = "activity.csv", header = TRUE)
head(activity)
str(activity)
summary(activity)

```

## What is mean total number of steps taken per day?

### Histogram with the total number of steps taken per day

```{r}
tot <- aggregate(steps~date, activity, sum)

library(lattice)
barchart(steps~date, data = tot)

```

### Table with the mean and median for the total number of steps taken per day

```{r}
ma <- aggregate(steps~date, activity, mean)
colnames(ma) <- c("date", "mean_steps")

me <- aggregate(steps~date, activity, median)
colnames(me) <- c("date", "median_steps")

merge(ma, me, by.x = "date", all.x = TRUE)
```

## What is the average daily activity pattern?

### Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days

```{r}
ts <- aggregate(steps~interval, activity, mean)
with(ts, plot(interval, steps, type = "l"))
```

### The 5-minute interval that contains the maximum number of steps is:

```{r}
maxx <- which(ts$steps == max(ts$steps))
ts[maxx,1]
```

## Imputing missing values

### Total number of rows with missing values are:

```{r}
comp <- complete.cases(activity)
activity_complete <- activity[comp,]
nrow(activity) - nrow(activity_complete)
```

### Filling in the missing values with the mean of that specific date

```{r}
new_activity <- activity

for (i in 1:17568){
  
  if(is.na(new_activity[i,1])){
    new_activity[i,1] <- ma[new_activity[i,2],2]
  }
}
head(new_activity)
```

### Now, with the new data set I will make a histogram of the total number of steps taken each day and calcula the mean and median to see what has changed

```{r}
tot_new <- aggregate(steps~date, new_activity, sum)

barchart(steps~date, data = tot_new)
```

We can see that the histogram differ from the previous one, when we counted the missing values.


## Are there differences in activity patterns between weekdays and weekends?

### Create a factor variable indicating if a given date is a week day or weekend

```{r}
week_d <- c("segunda-feira", "terça-feira", "quarta-feira", "quinta-feira", "sexta-feira")

for (i in 1:17568){
  
  if (weekdays(as.Date(new_activity[i,2])) %in% week_d){
    new_activity[i,4] <- "weekday"
  }
  else{
    new_activity[i,4] <- "weekend"
  }
}
colnames(new_activity) <- c("steps", "date", "interval", "weekday_or_weekend")
new_activity$weekday_or_weekend <- as.factor(new_activity$weekday_or_weekend)
head(new_activity)
```

### Plot the 5-minute interval and the average number of steps taken, averaged across all week days and weekends

```{r}
spl <- split(new_activity, new_activity$weekday_or_weekend)
activity_week <- spl[[1]]
activity_weekend <- spl[[2]]
```

For the week days:

```{r}
averg_week <- aggregate(steps~interval, activity_week, mean)
with(averg_week, plot(interval, steps, type = "l"))
```

For the weekends:

```{r}
averg_weekend <- aggregate(steps~interval, activity_weekend, mean)
with(averg_weekend, plot(interval, steps, type = "l"))
```
